---
title: "HW2"
output: github_document
---

loading library for hw2
```{r}
library(tidyverse)
```

Loading 3 datasets
```{r}
pols_data <-
  read_csv("~/Desktop/💻/按课程类型分类/P8105 Data Science 1/Homeworks/HW2/fivethirtyeight_datasets/pols-month.csv",na = c("NA",".","") ) |> 
  janitor::clean_names()

unemployment_data  <-
  read_csv("~/Desktop/💻/按课程类型分类/P8105 Data Science 1/Homeworks/HW2/fivethirtyeight_datasets/unemployment.csv",na = c("NA",".","") ) |> 
  janitor::clean_names()

snp_data  <-
  read_csv("~/Desktop/💻/按课程类型分类/P8105 Data Science 1/Homeworks/HW2/fivethirtyeight_datasets/snp.csv",na = c("NA",".","") )  |> 
  janitor::clean_names()
```

Use separate() to break up the variable mon into integer variables
```{r}
d <- pols_data |> 
  
  separate(mon, into = c("year","month","day"), sep = "-") |> 
  
    mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    day = as.numeric(day)
    ) |> 
  
  rename(gop = prez_gop) |> 
  rename(dem = prez_dem) |> 
  mutate(
    month = case_match(
      month,
      1 ~ "jan",
      2 ~ "feb",
      3  ~ "mar", 
      4  ~ "apr", 
      5  ~ "may", 
      6  ~ "jun", 
      7  ~ "jul", 
      8  ~ "aug", 
      9  ~ "sep", 
      10 ~ "oct", 
      11 ~ "nov",
      12 ~ "dec"
    ),
  month = as.factor(month)
  ) |> 
  select(-day) |> 
  pivot_longer(
    col = c(gop,dem),
    names_to = "president",
    values_to = "value"
  ) |> 
  select(-value) |> 
  arrange(desc(year))
d
```





```{r}
p <- snp_data |> 
  
  separate(date, into = c("month","day","year"), sep = "/") |>
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    day = as.numeric(day)
    ) |> 
  select(year,month,day,close) |> 
  mutate(
  year = if_else(year < 30, year + 2000, year + 1900)
) |> 
  mutate(
    month = case_match(
      month,
      1 ~ "jan",
      2 ~ "feb",
      3  ~ "mar", 
      4  ~ "apr", 
      5  ~ "may", 
      6  ~ "jun", 
      7  ~ "jul", 
      8  ~ "aug", 
      9  ~ "sep", 
      10 ~ "oct", 
      11 ~ "nov",
      12 ~ "dec"
    ),
  month = as.factor(month)
  ) |> 
  select(-day)|> 
  arrange(desc(year))

p
```




```{r}
q <- unemployment_data |> 
  pivot_longer(
    cols = c(jan, feb, mar, apr, may, jun, jul, aug, sep, oct, nov, dec),
    names_to = "month",
    values_to = "rate"
    )|> 
  arrange(desc(year))
q
```



```{r}
final_data <- d |> 
  left_join(q, by = c("year", "month")) |> 
  left_join(p, by = c("year", "month"))
final_data
```

The pols-month.csv data contains partisan control of the US presidential election and Congress, recorded on a monthly basis.

The snp.csv data contains closing prices for the S&P 500 index, documenting the performance of financial markets.

The unemployment.csv data provides the monthly unemployment rate for the US.

After cleaning and merging, we obtain a tidy data frame, `final_data`, with `r nrow(final_data)` rows and `r ncol(final_data)` columns, spanning approximately `r min(final_data$year)` to `r max(final_data$year)` years. Key variables include:

year, month: Time key

president: Presidential party (gov/dem)

close: S&P 500 monthly closing index

unemployment: Monthly unemployment rate

This merged dataset allows us to simultaneously examine the relationship between politics, the economy, and the labor market.



```{r}
library(readxl)
```

```{r}
Trash_Wheel_data <- 
  janitor::clean_names(read_excel("~/Desktop/💻/按课程类型分类/P8105 Data Science 1/Homeworks/HW2/202509 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel") ) 


Trash_Wheel_data <- Trash_Wheel_data |>
  mutate(year = as.numeric(year)) |> 
  select(
    "dumpster","month","year","date","weight_tons","volume_cubic_yards",
    "plastic_bottles","polystyrene","cigarette_butts","glass_bottles",
    "plastic_bags","wrappers","homes_powered","sports_balls"
  ) |> 
  filter(!is.na(date)) |> 
  mutate(sports_balls = as.integer( round(sports_balls ) ) ) |> 
  mutate(wheel_name   = "Mr. Trash Wheel" )
Trash_Wheel_data
```

```{r}
Professor_Trash_Wheel_data <- 
  janitor::clean_names(read_excel("~/Desktop/💻/按课程类型分类/P8105 Data Science 1/Homeworks/HW2/202509 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel") )

Professor_Trash_Wheel_data <- Professor_Trash_Wheel_data |> 
  select(
    "dumpster","month","year","date","weight_tons","volume_cubic_yards",
    "plastic_bottles","polystyrene","cigarette_butts","glass_bottles",
    "plastic_bags","wrappers","homes_powered"
  ) |> 
  filter(!is.na(date)) |> 
  mutate(wheel_name   = "Professor Trash Wheel" )
Professor_Trash_Wheel_data
```

```{r}
Gwynnda_Trash_Wheel_data <-
  janitor::clean_names(read_excel("~/Desktop/💻/按课程类型分类/P8105 Data Science 1/Homeworks/HW2/202509 Trash Wheel Collection Data.xlsx", sheet = "Gwynns Falls Trash Wheel") )

Gwynnda_Trash_Wheel_data <- Gwynnda_Trash_Wheel_data |> 
  select(
    "dumpster","month","year","date","weight_tons","volume_cubic_yards",
    "plastic_bottles","polystyrene","cigarette_butts",
    "plastic_bags","wrappers","homes_powered"
  ) |> 
  filter(!is.na(date)) |> 
  mutate(wheel_name   = "Gwynnda Trash Wheel" )
Gwynnda_Trash_Wheel_data
```

```{r}
Final_Trash_wheel_data <- bind_rows(Trash_Wheel_data,Professor_Trash_Wheel_data,Gwynnda_Trash_Wheel_data) |> 
  arrange(date, wheel_name)
Final_Trash_wheel_data
```

```{r}
Professor_total_weight <- Professor_Trash_Wheel_data |> 
  pull(weight_tons) |> 
  sum(na.rm = TRUE)
Professor_total_weight

Gwynnda_202206_cigarette <- Gwynnda_Trash_Wheel_data |> 
  filter(year(date) == 2022, month(date) == 6) |> 
  pull(cigarette_butts) |>
  sum(na.rm = TRUE)
Gwynnda_202206_cigarette
```


## Explaination
We collated and merged the datasets from the three collection devices, resulting in a tidy data frame with `r nrow(final_data)` observations. Examples of key variables include `r paste(head(final_data, 6), collapse = ", ")`.

Professor Trash Wheel collected a total of **`r round(Professor_total_weight, 1)` tons of trash in the available data.
Gwynnda collected a total of **`r scales::comma(Gwynnda_202206_cigarette)`** cigarette butts in **June 2022**.

For reproducibility, we imported the data using `readxl::read_excel()` with the specified worksheet and omitted rows and columns containing annotations/charts during cleaning (this was achieved by filtering out missing rows for `date` and `dumpster` and retaining only the relevant columns for trash items). Additionally, we round the *Sports Balls* variable to the nearest integer and convert it to integer type (`as.integer(round(...))`) to be consistent with downstream analyses.


```{r}
zip_data  <-
  read_csv("~/Desktop/💻/按课程类型分类/P8105 Data Science 1/Homeworks/HW2/zillow_data/Zip Codes.csv",na = c("NA",".","") ) |> 
  janitor::clean_names()

zip_data

zip_zori_data  <-
  read_csv("~/Desktop/💻/按课程类型分类/P8105 Data Science 1/Homeworks/HW2/zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv",na = c("NA",".","") ) |>
  janitor::clean_names()
zip_zori_data
```
```{r}
zip_data_1 <- zip_data |> 
  mutate(zip_code = as.character(zip_code)) |> 
  select(zip_code, county, neighborhood) |> 
  distinct() |> 
  arrange(zip_code, county) |> 
    distinct(zip_code, .keep_all = TRUE) |> 
  mutate(
    borough = case_when(
      county == "Bronx"    ~ "Bronx",
      county == "Kings"    ~ "Brooklyn",
      county == "New York" ~ "Manhattan",
      county == "Queens"   ~ "Queens",
      county == "Richmond" ~ "Staten Island",
    )
  )
zip_data_1
```

```{r}
zip_zori_data_1 <- zip_zori_data |> 
  pivot_longer(
    x2015_01_31:x2024_08_31,
    names_to = "date",
    values_to = "zori"
  ) |> 
  mutate(date = ymd(str_remove(date, "^x"))) |> 
  mutate(zip_code = as.character(region_name))
zip_zori_data_1
```

```{r}
final_tidy <- zip_zori_data_1 |> 
  left_join(zip_data_1, by = "zip_code") |> 

  select(
    zip_code, borough, neighborhood, date, zori,
    region_id, size_rank, region_type, state_name, state, city, metro, county_name
  ) |> 
  arrange(zip_code, date)
final_tidy
```
```{r}
x2020_2021_answer <- final_tidy |> 
  filter(
    (date >= ymd("2020-01-01") & date <= ymd("2020-01-31")) |
    (date >= ymd("2021-01-01") & date <= ymd("2021-01-31"))
  ) |> 
  mutate(date = format(date, "%Y-%m")) |> 
  select(zip_code, borough, neighborhood, date, zori) |> 
  pivot_wider(
    names_from = date, 
    values_from = zori
    ) |> 
  rename(Y2020 = `2020-01`, Y2021 = `2021-01`) |> 
  mutate(
    change = Y2021 - Y2020,
    change_percent = round(100 * change / Y2020 , 2)
  ) |> 
  arrange(change)

top10_change <- x2020_2021_answer |> 
  slice_head(n = 10) |> 
  select(zip_code, borough, neighborhood, Y2020, Y2021, change, change_percent)

top10_change
```
`r nrow(final_tidy)` observations, `r ncol(final_tidy)` variables， `r dplyr::n_distinct(final_tidy$zip_code)`distinct zip code, and `r dplyr::n_distinct(final_tidy$neighborhood, na.rm = TRUE)` distinct community

Before the merge, I removed the duplicated ZIP table, resolving rare duplicate rows caused by cross-county mis-entries, such as 11201 (should belong to Kings/Brooklyn) and 10463 (should belong to the Bronx). This ensured that after the merge, each ZIP had only one record per month.

Zillow's ZORI often excludes ZIPs designated for PO Boxes, work units, airports, and corporate buildings because they often lack available rental samples. A small number of "border/suburban ZIPs" (such as 115xx) may be included in the ZIP list, but they fall outside the coverage of the NYC five-county area or the Zillow NYC dataset and are therefore also missing.

ZORIs for some ZIPs in the early years (2015–2017) were missing, but coverage has increased annually. Therefore, when performing year-over-year/month-over-month comparisons, it is recommended to use fixed months and filter out missing data.